``FISTA'' in Banach spaces with adaptive discretisations

### 2022-01-25 response.tex

\newcommand*\justify{%
	\fontdimen2\font=0.4em% interword space
	\fontdimen3\font=0.2em% interword stretch
	\fontdimen4\font=0.1em% interword shrink
	\fontdimen7\font=0.1em% extra space
	\hyphenchar\font=`\-% allowing hyphenation
}
\newcommand{\review}[1]{\texttt{\justify{#1}}}
\newcommand{\todo}[1]{{\color{red}Todo: #1}}
\newcommand{\F}[1]{\mathbb{#1}}
\let\vecf\vec
\renewcommand{\vec}[1]{\mathbf{#1}}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\renewcommand{\ULthickness}{1.0pt}
\newcommand{\edit}[2]{{\ifmmode\text{\color{red}\sout{\ensuremath{#1}}}\else {\color{red} \sout{#1}}\fi} {\color{darkgreen} #2}}

\title{Response to reviewer comments:\\``FISTA'' in Banach spaces with adaptive discretisations}
\author{Antonin Chambolle and Robert Tovey}
\date{7th October 2021}

\begin{document}
\maketitle

We would like to thank both the reviewers again for their time and feedback. All changes are marked in the supplementary pdf and a detailed response to each reviewer is given below.

\section{Detailed response to first reviewer}
\review{If I had one more comment it would be that I find it somewhat odd the value of the paper's algorithm is presented ``for use with un-attained minima'' as the key theoretical novelty. As a point of critique: what does it mean to solve an optimization problem that doesn't have a solution? Do the solutions for the $U^n$-problems become meaningless asymptotically? Have the authors observed any odd behavior for large values of $n$? On the other hand, I think the authors sell themselves a bit short with the statement ``key theoretical novelty'' as the analysis could be useful for solving optimal control and PDE-constrained optimization problems with FISTA or related algorithms, where the interplay of mesh-size and algorithm convergence should always be taken into account.}

Thank you for the constructive feedback. That paragraph in the introduction has been re-worded to take into account some of your points. In an abstract sense one can always attain minima by adding points to the space (e.g. $\operatorname{argmin} e^{-u} = \{\infty\}$), as you say, the challenge is to have some regularity guarantees so that the iterates themselves are not `meaningless'. The meaning of `meaningless' depends on the topology. In the Lasso example, the iterates are meaningless in both $L^2$ and $L^1$, but can be understood in the space of measures (or distributions). In this sense, we have never observed anything unexpected for large $n$.

\section{Detailed response to second reviewer}

\begin{enumerate}
	\item \review{Do you assume that the norms of $H$, $U$ are related? I.e., do you assume that there is $c>0$ such that $\lVert u\rVert_U\leq c\lVert u\rVert_H$? Please clarify this in the paper.}

	There is no such requirement between the spaces, this is now clarified after the bullet points in Section 1.

	\item \review{(7) $u^*$ was not defined here}

	This has been clarified in the text, immediately above (7)

	\item \review{Lemma 6: what is $|||\cdot|||_*$?}

	In the statement of Lemma 6 we now clarify that this is the dual norm.

	\item \review{Definition 2: What is $q$? Smoothness of boundary is not needed. What is $h$? A given constant in $(0,1)$? The coefficients $\alpha,\beta$ have to be independent of $u_i^k$, and rather should depend only on $\omega_i^k$. There is a stray $u^*$ in the inequality defining order $p$. Also $U\subset L^q$ restricts $q$ to $[1,2]$.}

	Thank you for highlighting the errors in this definition. The values of $q$ and $h$ have been clarified. We have removed all mention of a particular basis, in response to your later comment, and clarified instead the general scaling properties (the $\alpha$ and $\beta$). You're right, the assumption $\F U\supset\F H$ was lazy and not what was intended. We only require $\F U$ to contain sub-levelsets of the energy, now clarified in the text. \todo{It now makes sense to consider $q>2$, even if Theorem 4 confirms that nothing is gained by doing so.}

	While it will be most common for $(\alpha,\beta)$ to depend only on $\omega$, it is not necessary. One could imagine an example where each sub-domain supports elements which are translations/rescaling/rotations of an original master element and the angular resolution of rotations also refines as $k$ increases. The only important factor is the volume rescaling, which we have tried to emphasise in Definition 2.

	\item \review{Theorem 4: I do not understand how Lemmas 14-16 proof the estimate of $a_U$.}

	You're absolutely right, Theorem 4 was not consistent with the lemmas in the appendix. We have kept the statement of Theorem 4 the same other than to remove the special case and remove the implicit assumption that $\tilde{\F U}^0$ was finite dimensional (previously part of Definition 2). We have also added a new lemma (Lemma 7) to improve clarity, it serves as a midpoint between Theorem 4 and Appendix B. The text in Appendix B now makes clear reference to Lemma 7 as part of the proof of Theorem 4.

	\item \review{(15): $A$ has to be continuous on $H$ as well.}

	Old (15) is now (18). This is now emphasised in two places: Lemma 8 is referenced to confirm that $A$ is continuous, and we require that $|A|^2\nabla f$ is 1-Lipschitz.

	\item \review{page 11, paragraph below (21): Sublevel sets of $E$ are not compact, they are not even bounded without further conditions on $f$ ($f$ bounded from below would be sufficient). How do you construct the minimizing sequence in $H$? How does $E(u^*_j) \to E(u^*)$ follow? I think an assumption is needed to ensure that $A$ is sequentially weak-star continuous, i.e., $A = B^*$ where $B : \mathbb{R}^m \to C([0, 1]^d)$.}

	As suggested, we now assume that $f$ is bounded from below and include a comment on weak-star continuity. The general argument has been re-worked to have an explicit sequence $\tilde w_k$ in $H$. We no longer rely on the existence of a minimising sequence (other than the explicit $\tilde w_k$).

	\item \review{section 6.2: $l^1\subset l^2$ in the countable case. The sentence ``then $u^*\in l^2$ makes the analysis simpler'' makes no sense.}

	This comment has been expanded upon in the text. The key point is the problem behaves much more like a finite-dimensional problem than infinite.

	\item \review{(25) What is $\partial_n E$?}

	Old (25) is now (27). This refers to the derivative of $E$ on the subspace $\F U^n$, this is now clarified immediately above it in the text.

	\item \review{Inequalities (61),(62) do not make sense: the quantities on the left and right-hand sides are equal.}

	It's not quite clear which equations are being referred to here. On the copy I see, the old (61),(62) are now (63),(64) which are equalities. The best match I can see are the current equations (61),(62) in Lemma 9. The first equation is just H\"older's inequality, there is no reason for it to be equality for general $r$. The second inequality comes from the equivalence of norms, $\frac{\lVert\cdot\rVert_q}{\lVert\cdot\rVert_2}$.

	\todo{I don't understand this. (61),(62) are statements of Fourier/Gaussian data in numerics section, his next point is also on page 18 rather than 17. (51),(52) are in the long align environment in section 6.4 (support detection), those are plausible. (41),(42) are in section 6.3.2 (bounds for countable functions). (59),(60) are the two inequalities of Lemma 9. My best guess is this last pair (which are ironically now (61),(62)...). This is quite a patronising answer given the general standard of his questions thought.}

	\item \review{page 17, last sentence of first paragraph: the problem is convex, hence all critical points are globally optimal.}

	You're quite right, we meant criticality in the discrete sense. This has been corrected.

	\item \review{Lemma 14: domains are open sets, so ``compact domain'' does not make sense. The assumptions already restrict $q$ to $[1, 2]$, so $q > 2$ is impossible. Clash of notations: $\mathbb U$ is used to denote a finite-dimensional subspace of $\mathbb H \subset \mathbb U$. How is the assumption $e_j \in \mathbb U^*$ meant? Quantifier ``$\forall j$'' is missing in items 2,3. The proof uses orthonormality of the basis vectors, this assumption is not mentioned in Section 5. It is also rather unconventional in the context of finite elements. The last sentence of the proof seems to be not belong to the proof.}

	The old Lemma 14 is now Lemma 15. The domain is now assumed only to be bounded and $\tilde{\F{U}}$ is used as the subspace of $\F U$. We have removed the reference to orthonormal bases and the space $\F{U}^*$ which leads to a shorter and simpler result.

	\item \review{Lemma 15: What does ``value of $C$ satisfies the conditions of Lemma 14'' mean? The proof of Lemma 15 seems to use the proof of Lemma 14 but not the result of Lemma 14. It should be $e_j$ in (115) instead of $u_j$ . Also what does ``we compute the scaling constant'' mean? The estimate (116) has to be performed for all elements of $\mathbb U^k$ not only for basis elements.}

	The old Lemma 15 is now Lemma 16. Again, you were quite right, we now make explicit use of Lemma 15. This combined with the removal of orthonormal bases leaves a much more direct proof.

	\item \review{Lemma 16: Exponent $k$ is missing in all estimates. What is $u^*$? How is the first inequality in (119) obtained?}

	The old Lemma 16 is now Lemma 17. We have clarified that $u^*$ is a minimiser of $E$. The third case has been removed, so the old (119) has also been removed.
\end{enumerate}

### 2022-01-25 paper.tex
Second revision, first draft

\usepackage{amsmath,enumitem,datetime,xfrac,fix-cm,color,multirow,siunitx}
\usepackage{amssymb,mathtools,stmaryrd,dsfont,tikz}
\usepackage{fullpage}
\usepackage[colorlinks, citecolor=blue, bookmarks=True]{hyperref}
\usetikzlibrary{positioning,trees}
\usepackage{algorithm,algpseudocode}
%\graphicspath{{Figures/}}
\DeclareGraphicsExtensions{.pdf,.png}

\let\F\mathds\let\C\mathcal\newcommand{\R}{\F{R}}\newcommand{\A}{\tens{A}}
\newcommand{\norm}[1]{{\left\lVert #1 \right\rVert}}
\newcommand{\Norm}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}
\newcommand{\IP}[2]{\left\langle #1,#2 \right\rangle}
\newcommand{\ip}[2]{#1 \vcenter{\hbox{\resizebox{6pt}{!}{\ensuremath\cdot}}} #2}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\splitln}[4]{\begin{cases} #1 & #2 \\ #3 & #4\end{cases}}
\newcommand{\1}{\F{1}}
\DeclareMathOperator{\st}{\;s.t.\;}
\renewcommand{\epsilon}{\varepsilon}\renewcommand{\bar}{\overline}
\renewcommand{\hat}{\widehat}\renewcommand{\tilde}{\widetilde}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sign}{sign}
\newcommand{\Emin}[1][\var0]{\inf_{#1\in\F H}\op{E}(#1)}

@@

% Macro for tracking changes
\usepackage[normalem]{ulem}\renewcommand{\ULthickness}{1.0pt}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\newcommand{\edit}[2]{{\ifmmode\text{\color{red}\sout{\ensuremath{#1}}}\else {\color{red} \sout{#1}}\fi} {\color{darkgreen} #2}}
\renewcommand{\edit}[2]{#2}
%\newcommand{\todo}[1]{{\color{red}#1}}

@@

	\begin{abstract}
		FISTA is a popular convex optimisation algorithm which is known to converge at an optimal rate whenever \edit{the optimisation domain}{a minimiser} is contained in a suitable Hilbert space. We propose a modified algorithm where each iteration is performed in a \edit{subspace, and that subspace is allowed to change at every iteration}{subset which is allowed to change at every iteration}. \edit{Analytically, this allows us to guarantee convergence in a Banach space setting}{Sufficient conditions are provided for guaranteed convergence}, although at a reduced rate depending on the conditioning of the specific problem. \edit{Numerically we show that a greedy adaptive choice of discretisation can greatly increase the time and memory efficiency in infinite dimensional Lasso optimisation problems.}{These conditions have a natural interpretation when a minimiser exists in an underlying Banach space. Typical examples are L1-penalised reconstructions where we provide detailed theoretical and numerical analysis.}
	\end{abstract}
	\keywords{Convex optimization \and Multiscale \and Multigrid \and Sparsity \and Lasso}

	\section{Introduction}
	The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was proposed by Beck and Teboulle \cite{Beck2009} as an extension of Nesterov's fast gradient method \cite{Nesterov2004} and is now a very popular algorithm for minimising the sum of two convex functions.
	We write this as the problem of computing
	\begin{equation}\label{eq: min F = fD+fC}
		\edit{u^* \in \argmin_{\var0\in\F H}\op{E}(\var0)}{\Emin}  \qquad\text{such that}\qquad \op{E}(\var0)\coloneqq \op{f}(\var0) + \op{g}(\var0),
	\end{equation}
	where $\op{f}\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\op{g}\colon\F H\to\bar\R$ is a convex function, whose ``proximity operator'' is easy to compute. The iterates of the FISTA algorithm will be denoted $\var0_n\in\F H$. It has been shown that $\op{E}(\var0_n) - \edit{\op{E}(u^*)}{\Emin}$ converges at the optimum rate of $n^{-2}$ \cite{Beck2009}, and later (after a small modification) the convergence of the iterates was also shown in a general Hilbert space setting \cite{Chambolle2015}. Many further works have gone on to demonstrate faster practical convergence rates for slightly modified variants of FISTA \cite{Tao2016,Liang2017,Alamo2019}.

	The contribution of this work is a modified FISTA algorithm which achieves two key properties:
	\begin{itemize}
		\item we extend the convergence results to the case \edit{$\var0^*\in\bar{\F H}\setminus\F H$ where $\F{U}\coloneqq \bar{\F H}$ is a Banach space}{where the minimum is not achieved. For example, in \edit{Lasso}{L1 penalised reconstructions} discussed in Section~\ref{sec: Lasso definition} there exists a Banach space $\F{U}\supset\F H$ such that $\Emin = \min_{\var0\in\F{U}}\op{E}(\var0)$ is achieved in $\F{U}\setminus\F H$}. The proven rate is slower than the optimal $n^{-2}$, but appears to be quite sharp in numerical examples.
		\item we allow for adaptive discretisation such that $\var0_n\in\F{U}^n\subset\F H$. With simple guarantees on the rates of convergence, we show analytically that applying our FISTA algorithm with inexact discretisation does not reduce the proven rate of convergence. In the Hilbert space setting this is the original $n^{-2}$ rate. In numerical examples we demonstrate that greedy adaptive discretisation strategies can achieve much greater performance than uniform discretisation, both in measures of time and computer memory requirements.
	\end{itemize}
	There is much overlap between the techniques used in this work and those used in the literature of inexact optimisation, although we believe that our interpretation for \edit{the use outside of Hilbert spaces}{use with un-attained minima} is the key theoretical novelty. Some other works for FISTA-like algorithms include \cite{Jiang2012,Villa2013}. Of particular note, our stability estimate for FISTA  in Theorem~\ref{thm: mini FISTA convergence} is very similar to \cite[Prop 2]{Schmidt2011} and \cite[Prop 3.3]{Aujol2015}. This is then used to analyse the convergence properties in our more general Banach space setting where all sources of inexactness come from subspace approximations. The ideas in \cite{Parpas2017} are similar although in application to the proximal gradient method with an additional smoothing on the functional $\op{g}$. The permitted refinement steps are also more broad in our work. \cite{}{Very recent work in \cite{Yu2021} proposes a `Multilevel FISTA' algorithm which allows similar coarse-to-fine refinement strategies, although only a finite number. We also allow for non-uniform refinement with a posteriori strategies.}

	\subsection{Outline}
	This work is organised as follows. Section~\ref{sec: prelims} defines notation and the generic form of our proposed refining FISTA algorithm, Algorithm~\ref{alg: refining FISTA}. The main theoretical contribution of this work is the convergence analysis of Algorithm~\ref{alg: refining FISTA} which is split into two parts: first we outline the proof structure in Section~\ref{sec: recipe}, then we state the specific results in the case of FISTA in Section~\ref{sec: FISTA convergence}. The main results are Theorems~\ref{thm: exponential FISTA convergence}/\ref{thm: stronger exponential FISTA convergence} which extend the convergence of FISTA \edit{to infinite dimensional Banach spaces}{to cases with un-attained minima} with uniform/adaptively chosen subspaces $\F{U}^n$ respectively.

	Section~\ref{sec: general examples} presents some general results for the application of Algorithm~\ref{alg: refining FISTA} \edit{}{in Banach spaces }and Section~\ref{sec: Lasso definition} gives a much more detailed discussion of adaptive refinement for Lasso minimisation. In particular, we describe how to choose efficient refining discretisations to approximate \edit{$\var0^*$}{$\Emin$}, estimate the convergence of $\op{E}$, and identify the support of \edit{$\var0^*$}{the minimiser}. The numerical results in Section~\ref{sec: numerics} demonstrate these techniques in four different models demonstrating the apparent sharpness of our convergence rates and the computational efficiency of adaptive discretisations.

	\section{Definitions and notation}\label{sec: prelims}
	\edit{We consider optimisation of \eqref{eq: min F = fD+fC} over two spaces, the Banach space $(\F{U},\Norm\cdot)$ and Hilbert space $(\F H,\IP\cdot\cdot,\norm\cdot)$, such that 
	\\\centerline{$\displaystyle \exists\var0^*\in \F{U} \quad\st\quad \op{E}(\var0^*) = \min_{\var0\in\F{U}}\op{E}(\var0) = \inf_{\var0\in\F H}\op{E}(\var0).$}
	Furthermore, define the translated energy
	\\\centerline{$\displaystyle \op{E}_0\colon\F{U}\cup\F H\to \R,\qquad \op{E}_0(\var0) \coloneqq \op{E}(\var0)-\op{E}(\var0^*).$}
	Note that the value of $\op{E}_0$ is uniquely defined, even if $\var0^*$ may not be.%
	\\}{%
	We consider optimisation of \eqref{eq: min F = fD+fC} over a Hilbert space $(\F H,\IP\cdot\cdot,\norm\cdot)$ with the translated energy
	\begin{equation}
		\op{E}_0\colon\F H\to \R,\qquad \op{E}_0(\var0) \coloneqq \op{E}(\var0)-\Emin[\tilde{\var0}].
	\end{equation}
	}

	The proposed generalised FISTA algorithm is stated in Algorithm~\ref{alg: refining FISTA} for an arbitrary choice of \edit{refining subspaces}{closed convex subsets} $\F{U}^n\subset \edit{\F{U}\cap}{}\F H$ for $n\in\F N$. The only difference \edit{}{from standard FISTA }is that on iteration $n$, all computations are performed in the \edit{subspace}{subset} $\F{U}^n$. If $\F{U}^n\edit{=\F{U}}{}=\F H$, then we recover the \edit{standard FISTA}{original} algorithm. \edit{}{More generally, the idea is that $\F{U}^n$ are `growing', for example $\F{U}^n\subset\F{U}^{n+1}$, but this assumption is not necessary in most of the results.}

	Without loss of generality we will assume $L=1$, i.e. $\nabla\op{f}$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\op{E}$ with $\frac{\op{E}}{L}$.\edit{ A more important distinction is to say that the properties of $\op{f}$ and $\op{g}$ are stated with respect to the Hilbert norm.}{} In particular,
	\begin{equation}
		\norm{\nabla\op{f}(\var0)- \nabla\op{f}(\var1)} \leq \norm{\var0-\var1}
	\end{equation}
	for all $\var0,\var1\in\F H$ and $\op{g}$ is called `simple' if \edit{}{it is proper, convex, weakly lower-semicontinuous, and}
	\begin{equation}
		\argmin_{\var0\in\tilde{\F{U}}}\tfrac12\norm{\var0-\var1}^2 + \op{g}(\var0)
	\end{equation}
	is exactly computable for all $\var1\in\F H$ and all $\tilde{\F{U}}\in\{\F{U}^n\}_{n=0}^\infty$. \edit{}{Closed subsets of $\F H$ are locally weakly compact, therefore this argmin is always non-empty.}

@@

	\edit{Finally we introduce the concept of an orthogonal projection in the setting of this work. For a subspace $\F{U}^n\subset\F{U}\cap\F H$ we define the orthogonal projection $\tens{\Pi}_n\colon (\F{U}^n)^*\to \F{U}^n$ to be any extension of the function such that 
	\\\centerline{$\IP{\tens{\Pi}_n\var0}{\var1} = \IP{\var0}{\var1} \text{ for all } \var0\in(\F{U}^n)^*, \ \var1\in\F{U}^n.$}
	This definition is non-standard as we view $(\F{U}^n)^*$ embedded in $\F{U}$. Note that $\F{U}^n\subset\F H$ implies that $(\F{U}^n)^*\supset\F H$, therefore $\tens{\Pi}_n$ is an extension of the classical orthogonal projection. It is convenient to allow an extension here because $\F{U}$ may be bigger than $\F H$, $(\F{U}^n)^*$ is the largest space such that $\tens{\Pi}_n$ is sufficiently well-defined. Beyond this, the Hahn-Banach theorem may allow the domain of $\tens{\Pi}_n$ to be extended further, but the definition is no longer unique. For example, in 1D the value of $\IP{\delta_0}{\frac{1}{\sqrt[4]{x}}}$ is not well-defined, but the value of $\IP{\delta_0}{\1_{[0,1]}}$ can be chosen to be consistent with some sequence of molifiers. To account for this possibility, we shall state that the domain of $\tens{\Pi}_n$ is $\bar{(\F{U}^n)^*}$, with the closure computed in an appropriate topology dictated by $\Norm\cdot$.}{}

>>>>

\usepackage{amsmath,enumitem,datetime,xfrac,fix-cm,color,multirow,siunitx}
\usepackage{amssymb,mathtools,stmaryrd,dsfont,tikz}
\usepackage{fullpage}
\usepackage[colorlinks, citecolor=blue, bookmarks=True]{hyperref}
\usetikzlibrary{positioning,trees}
\usepackage{algorithm,algpseudocode}
\graphicspath{{Figures/}}
\DeclareGraphicsExtensions{.pdf,.png}

\let\F\mathds\let\C\mathcal\newcommand{\R}{\F{R}}\newcommand{\A}{\tens{A}}
\newcommand{\norm}[1]{{\left\lVert #1 \right\rVert}}
\newcommand{\Norm}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}
\newcommand{\IP}[2]{\left\langle #1,#2 \right\rangle}
\newcommand{\ip}[2]{#1 \vcenter{\hbox{\resizebox{6pt}{!}{\ensuremath\cdot}}} #2}
\newcommand{\op}[1]{\operatorname{#1}}
\newcommand{\splitln}[4]{\begin{cases} #1 & #2 \\ #3 & #4\end{cases}}
\newcommand{\1}{\F{1}}
\DeclareMathOperator{\st}{\;s.t.\;}
\renewcommand{\epsilon}{\varepsilon}\renewcommand{\bar}{\overline}
\renewcommand{\hat}{\widehat}\renewcommand{\tilde}{\widetilde}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\sign}{sign}
\newcommand{\diff}{\mathop{}\!\mathrm{d}}
\newcommand{\Emin}[1][\var0]{\inf_{#1\in\F H}\op{E}(#1)}
\newcommand{\aU}{a_{\op{U}}}\newcommand{\aE}{a_{\op{E}}}
\newcommand{\aUaE}{$(\aU,\aE)$-minimising sequence of $\op{E}$}

@@

% Macro for tracking changes
\usepackage[normalem]{ulem}\renewcommand{\ULthickness}{1.0pt}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\newcommand{\edit}[3][1]{
	\IfEq{#1}{2}{\def\mysecondvar{}}{\def\mysecondvar{#2}}% hide the deleted text?
	\ifx\mysecondvar\empty{}\else{%
		\IfEq{#1}{1}{%
			\ifmmode%
			\text{\color{red}\sout{\ensuremath{#2}}}%
			\else%
			{\color{red} \sout{#2}}%
			\fi%
		}{\color{red}#2}\ %
	}\fi{\color{darkgreen}#3}}
%\renewcommand{\edit}[3][1]{#3}
\newcommand{\todo}[1]{{\color{orange}#1}}

@@

	\begin{abstract}
		FISTA is a popular convex optimisation algorithm which is known to converge at an optimal rate whenever a minimiser is contained in a suitable Hilbert space. We propose a modified algorithm where each iteration is performed in a subset which is allowed to change at every iteration. Sufficient conditions are provided for guaranteed convergence, although at a reduced rate depending on the conditioning of the specific problem. These conditions have a natural interpretation when a minimiser exists in an underlying Banach space. Typical examples are L1-penalised reconstructions where we provide detailed theoretical and numerical analysis.
	\end{abstract}
	\keywords{Convex optimization \and Multiscale \and Multigrid \and Sparsity \and Lasso}

	\section{Introduction}
	The Fast Iterative Shrinkage-Thresholding Algorithm (FISTA) was proposed by Beck and Teboulle \cite{Beck2009} as an extension of Nesterov's fast gradient method \cite{Nesterov2004} and is now a very popular algorithm for minimising the sum of two convex functions.
	We write this as the problem of computing
	\begin{equation}\label{eq: min F = fD+fC}
		\Emin \qquad\text{such that}\qquad \op{E}(\var0)\coloneqq \op{f}(\var0) + \op{g}(\var0),
	\end{equation}
	where $\op{f}\colon\F H\to\R$ is a convex differentiable function with $L$-Lipschitz gradient and $\op{g}\colon\F H\to\bar\R$ is a convex function, whose ``proximity operator'' is easy to compute. The iterates of the FISTA algorithm will be denoted $\var0_n\in\F H$. It has been shown that $\op{E}(\var0_n) - \Emin$ converges at the optimum rate of $n^{-2}$ \cite{Beck2009}, and later (after a small modification) the convergence of the iterates was also shown in a general Hilbert space setting \cite{Chambolle2015}. Many further works have gone on to demonstrate faster practical convergence rates for slightly modified variants of FISTA \cite{Tao2016,Liang2017,Alamo2019}.

\edit[0]{%
	\sout{The contribution of this work is a modified FISTA algorithm which achieves two key properties:}
	\begin{itemize}
		\item \sout{We extend the convergence results to the case where the minimum is not achieved. For example, in L1 penalised reconstructions discussed in Section~\ref{sec: Lasso definition} there exists a Banach space $\F{U}\supset\F H$ such that $\Emin = \min_{\var0\in\F{U}}\op{E}(\var0)$ is achieved in $\F{U}\setminus\F H$. The proven rate is slower than the optimal $n^{-2}$, but appears to be quite sharp in numerical examples.}
		\item \sout{We allow for adaptive discretisation such that $\var0_n\in\F{U}^n\subset\F H$. With simple guarantees on the rates of convergence, we show analytically that applying our FISTA algorithm with inexact discretisation does not reduce the proven rate of convergence. In the Hilbert space setting this is the original $n^{-2}$ rate. In numerical examples we demonstrate that greedy adaptive discretisation strategies can achieve much greater performance than uniform discretisation, both in measures of time and computer memory requirements.}
	\end{itemize}
	\sout{There is much overlap between the techniques used in this work and those used in the literature of inexact optimisation, although we believe that our interpretation for use with un-attained minima is the key theoretical novelty.}

}{%
	There is much overlap between the techniques used in this work and those used in the literature of inexact optimisation, however, our interpretation is relatively novel. In particular, we emphasise the infinite-dimensional setting where errors come from `discretisation' (rather than random or decaying errors in $\F H$), which enables two new perspectives:
	\begin{itemize}
		\item Analytically, we prove new rates of convergence for FISTA when the minimum energy is not achieved (at least not in $\F H$). The exact rate can be computed by quantifying coercivity and regularity properties of $\op{E}$. If there isn't a minimiser in $\F H$, then this rate is strictly smaller than $n^{-2}$.

		\item Numerically, we allow the optimisation domain to change on every iteration. This enables us to understand how FISTA behaves with adaptive discretisations. Adaptive finite-element methods are known to improve the efficiency of, for example, approximating the solutions of PDEs. Our analytical results show how to combine such tools with FISTA (without reducing the guaranteed rate of convergence), and our numerical results confirm much improved time and computer memory efficiency in the Lasso example (Section~\ref{sec: Lasso definition}).
	\end{itemize}
	All the examples in this work, discussed from Section~\ref{sec: general examples} onward, consider $\{\var0\in\F H\st \op{E}(\var0)<\infty\}$ to be contained in some ambient Banach space $\F{U}$. The idea is that FISTA provides a minimising sequence in $\F H\cap\F U$, but further properties like rate of convergence (of $\op{E}$ or iterates) must come from the topology of $\F{U}$. It will not be necessary for $\F H\hookrightarrow\F U$ to be a continuous embedding, nor in fact the full inclusion $\F H\subset\F U$.

	\todo{Note to self: It is probably not very hard to imagine examples outside of a Banach space, you just need to find a metric space without the linear scaling. Examples outside of a metric space would be much harder.}
}

	Some other works for FISTA-like algorithms include \cite{Jiang2012,Villa2013}. Of particular note, our stability estimate for FISTA  in Theorem~\ref{thm: mini FISTA convergence} is very similar to \cite[Prop 2]{Schmidt2011} and \cite[Prop 3.3]{Aujol2015}. This is then used to analyse the convergence properties in our more general Banach space setting where all sources of inexactness come from subspace approximations. The ideas in \cite{Parpas2017} are similar although in application to the proximal gradient method with an additional smoothing on the functional $\op{g}$. The permitted refinement steps are also more broad in our work. Very recent work in \cite{Yu2021} proposes a `Multilevel FISTA' algorithm which allows similar coarse-to-fine refinement strategies, although only a finite number. We also allow for non-uniform refinement with a posteriori strategies.

	\subsection{Outline}
	This work is organised as follows. Section~\ref{sec: prelims} defines notation and the generic form of our proposed refining FISTA algorithm, Algorithm~\ref{alg: refining FISTA}. The main theoretical contribution of this work is the convergence analysis of Algorithm~\ref{alg: refining FISTA} which is split into two parts: first we outline the proof structure in Section~\ref{sec: recipe}, then we state the specific results in the case of FISTA in Section~\ref{sec: FISTA convergence}. The main results are Theorems~\ref{thm: exponential FISTA convergence}/\ref{thm: stronger exponential FISTA convergence} which extend the convergence of FISTA to cases with un-attained minima with uniform/adaptively chosen subspaces $\F{U}^n$ respectively.

	Section~\ref{sec: general examples} presents some general results for the application of Algorithm~\ref{alg: refining FISTA} in Banach spaces and Section~\ref{sec: Lasso definition} gives a much more detailed discussion of adaptive refinement for Lasso minimisation. In particular, we describe how to choose efficient refining discretisations to approximate $\Emin$, estimate the convergence of $\op{E}$, and identify the support of the minimiser. The numerical results in Section~\ref{sec: numerics} demonstrate these techniques in four different models demonstrating the apparent sharpness of our convergence rates and the computational efficiency of adaptive discretisations.

	\section{Definitions and notation}\label{sec: prelims}
	We consider optimisation of \eqref{eq: min F = fD+fC} over a Hilbert space $(\F H,\IP\cdot\cdot,\norm\cdot)$ with the translated energy
	\begin{equation}
		\op{E}_0\colon\F H\to \R,\qquad \op{E}_0(\var0) \coloneqq \op{E}(\var0)-\Emin[\tilde{\var0}].
	\end{equation}

	The proposed generalised FISTA algorithm is stated in Algorithm~\ref{alg: refining FISTA} for an arbitrary choice of closed convex subsets $\F{U}^n\subset \F H$ for $n\in\F N$. The only difference from standard FISTA is that on iteration $n$, all computations are performed in the subset $\F{U}^n$. If $\F{U}^n=\F H$, then we recover the original algorithm. More generally, the idea is that $\F{U}^n$ are `growing', for example $\F{U}^n\subset\F{U}^{n+1}$, but this assumption is not necessary in most of the results.

	Without loss of generality we will assume $L=1$, i.e. $\nabla\op{f}$ is 1-Lipschitz. To get the general statement of any of the results which follow, replace $\op{E}$ with $\frac{\op{E}}{L}$. In particular,
	\begin{equation}
		\norm{\nabla\op{f}(\var0)- \nabla\op{f}(\var1)} \leq \norm{\var0-\var1}
	\end{equation}
	for all $\var0,\var1\in\F H$ and $\op{g}$ is called `simple' if it is proper, convex, weakly lower-semicontinuous, and
	\begin{equation}
		\argmin_{\var0\in\tilde{\F{U}}}\tfrac12\norm{\var0-\var1}^2 + \op{g}(\var0)
	\end{equation}
	is exactly computable for all $\var1\in\F H$ and all $\tilde{\F{U}}\in\{\F{U}^n\}_{n=0}^\infty$. Closed subsets of $\F H$ are locally weakly compact, therefore this argmin is always non-empty.

@@

	The precise constants associated to a given rate are given in the statements of the theorems but, for convenience, are otherwise omitted from the text. For sequences $(a_n)_{n=1}^\infty$,$(b_n)_{n=1}^\infty$ we will use the notation:
	\begin{align*}
		a_n \lesssim b_n \qquad &\iff \qquad \exists C,N>0 \st a_n \leq C b_n \text{ for all } n>N,
		\\ a_n \simeq b_n \qquad &\iff \qquad a_n\lesssim b_n \lesssim a_n.
	\end{align*}
	For $n\in\F N$ we use the abbreviation $[n] = \{1,2,\ldots,n\}$. When the subdifferential of $\op{E}$ is set-valued, we will use the short-hand
	\begin{equation}
		\Norm{\partial\op{E}(\var0)} \coloneqq \inf_{\var1\in\partial\op{E}(\var0)}\Norm{\var1}
	\end{equation}
	for any specified norm $\Norm\cdot$.

	\begin{algorithm}\caption{Refining subset FISTA}\label{alg: refining FISTA}
		\centering
		\begin{algorithmic}[1]
			\State Choose $(\F{U}^n)_{n\in\F N}$, $\var0_0\in \F{U}^0$ and some FISTA stepsize choice $(t_n)_{n\in\F N}$
			\State $\var1_0\gets \var0_0, n\gets0$
			\Repeat
			\State $\bar{\var0}_n \gets (1-\tfrac{1}{t_n})\var0_n + \tfrac{1}{t_n}\var1_n$
			\State $\displaystyle \var0_{n+1} \gets \argmin_{\var0\in \F{U}^{n+1}} \tfrac12\norm{\var0-\bar{\var0}_n+\nabla \op{f}(\bar{\var0}_n)}^2+\op{g}(\var0)$ \Comment{Only modification, $\F{U}^{n+1}\subset \F{U}$}
			\State $\var1_{n+1} \gets (1-t_n)\var0_n + t_n\var0_{n+1}$
			\State $n\gets n+1$
			\Until{converged}
		\end{algorithmic}
	\end{algorithm}
